---
title: "What Decides Wine Quality"
author: "Allen Martinez, Mingjun Li, Yifei Zhou"
date: "9/23/2019"
output:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r echo=FALSE}
wine <- read.csv('winedata.csv',stringsAsFactors = F)
attach(wine)
```
# Introduction
Background:

Purpose: 

- relationship between wine quality and a series of chemical composition indexes of wine
- give some advice on how to improve the wine procedure


# Data Overview
- detect missing value
```{r echo=FALSE}
table(is.na(wine))
```
- data structure
```{r wine, echo = TRUE}
str(wine)
```


# Variables Explaination


# Data distribution (1)
```{r echo=FALSE}
library(ggplot2)
par(mfrow=c(2,3))
hist(quality_score)
for (i in 2:6){
  boxplot(wine[,i]~type,ylab=names(wine)[i],xlab='',names=c('white','red'))
}
```


# Data distribution (2)
```{r echo=FALSE}
par(mfrow=c(2,3))
for (i in 7:12){
  boxplot(wine[,i]~type,ylab=names(wine)[i],xlab='',names=c('white','red'))
}
```


# Relationship with Quality Score (1)
```{r echo=FALSE}
library(ggplot2)
hist(type,breaks=2,xlab='Type of Wines (0 for white wine and 1 for red wine)')
for (i in 2:6){
  plot = ggplot(wine,aes(x=wine[,i],y=quality_score,col=factor(type)))+
        geom_point() +
        scale_color_manual('0 for white, 1 for red',values=c('skyblue','red3'))+
        labs(x=names(wine)[i])
  print(plot)
}
```


# Relationship with Quality Score (2)
```{r echo=FALSE}
library(ggplot2)
for (i in 7:12){
  plot = ggplot(wine,aes(x=wine[,i],y=quality_score,col=factor(type)))+
        geom_point() +
        scale_color_manual('0 for white, 1 for red',values=c('skyblue','red3'))+
        labs(x=names(wine)[i])
  print(plot)
}

```


# Regression Procedures 



# Fit Mulitple Linear Regression Model With All Variables
```{r echo=FALSE}
m1 <- lm(quality_score ~ type+fixed.acidity+volatile.acidity+citric.acid+residual.sugar+
           chlorides+free.sulfur.dioxide+total.sulfur.dioxide+wine$density+pH+sulphates+alcohol)
summary(m1)
```

# Diagnostic Plots
```{r echo=FALSE}
par(mfrow=c(2,2))
plot(m1)
```

# Challenges
- Unconstant Variance --> Box Cox

- Multicollinearity --> Delete Variables

- Influential Plot(tend to overestimate) --> Reweight


# Box Cox
>best lamba = 0.7475

```{r echo=FALSE,results='hide'}
library(MASS)
bc <- boxcox(m1)
lambda <- bc$x 
likelihood <- bc$y 
bestlambda <- lambda[which(likelihood==max(likelihood))]
quality_score.t = quality_score**(bestlambda)
```
```{r echo=FALSE}
m2 <- lm(quality_score.t ~ type+fixed.acidity+volatile.acidity+citric.acid+residual.sugar+
          chlorides+free.sulfur.dioxide+total.sulfur.dioxide+density+pH+sulphates+alcohol)
par(mfrow=c(2,2))
plot(m2)
```

# After Transformation
> Current model has serious Multicollinearity

```{r echo=FALSE}
library(car)
```
```{r echo=FALSE}
vif(m2)
```

> Delete the density

```{r echo=FALSE}
m4 <- lm(quality_score.t ~ type+fixed.acidity+volatile.acidity+citric.acid+residual.sugar+
          chlorides+free.sulfur.dioxide+total.sulfur.dioxide+pH+sulphates+alcohol)
vif(m4)
```
```{r echo=FALSE,results='hide'}
summary(m4)
```

# Variable Selection
# Type 3 Extra SSE
> fixed.acidity, citric.acid and pH do not significantly help to explain the data.

```{r echo=FALSE}
library(car)
Anova(m4,type=3)
```

# Best Subset
```{r echo=FALSE}
library(leaps)
bestsubsets <- regsubsets(
          quality_score.t ~ type+fixed.acidity+volatile.acidity+citric.acid+residual.sugar+
          chlorides+free.sulfur.dioxide+total.sulfur.dioxide+pH+sulphates+alcohol,
          data=wine,nbest=1,nvmax = 11)
s <- summary(bestsubsets)
cbind(s$which,r2=s$rsq,adjr2=s$adjr2,cp=s$cp,bic=s$bic)
```


# Stepwise
> forward result: 9 vars

```{r echo=FALSE,results='hide'}
null <- lm(quality_score.t~1)
full <- m4
forward <- step(null,scope=list(lower=null,upper=full),direction = 'both')
```
```{r echo=FALSE}
summary(forward)
```
# Stepwise
> backward result: same as forward, R_adj is small

```{r echo=FALSE,results='hide'}
backward <- step(full,direction='both')
```
```{r echo=FALSE}
summary(backward)
```


# Detecing Influential Points
```{r echo=FALSE}
library(car)
best <- lm(quality_score.t ~ type + volatile.acidity + citric.acid + 
    residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + 
    sulphates + alcohol)
summary(best)

```
```{r}
par(mfrow=c(2,2))
plot(best)
```
```{r echo=FALSE}
fitted <- best$fitted
leverage = hatvalues(best)

par(mfrow=c(2,2))
# High leverage points
threshold = 2*length(best$coefficients)/length(quality_score)
plot(fitted,leverage,main='Plot of Leverage',type='n')
text(fitted,leverage,labels=round(quality_score,0))
abline(h=threshold)

# Influential Plot
influencePlot(best,main='Influence Plot')
cook <- cooks.distance(best)
# Cook distance Plot
plot(fitted,cook,main='Plot of Cook Distance',type='n')
text(fitted,cook,labels=round(quality_score,0))
```
```{r echo=FALSE,results='hide'}
# DFITS
Infval <- dffits(best)
subset(Infval, Infval > 2*sqrt(length(best$coefficients)/length(quality_score)))
```

# Influential Data Points
```{r echo=FALSE,result='asis'}
# show the influential points
wine[c(1206, 1908, 4720, 5129, 5022),c("type","quality_score")]
```

# Reweight Regression Model
```{r echo=FALSE}
library(MASS)
rlm <- rlm( quality_score.t ~ type + volatile.acidity + citric.acid + 
    residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + 
    sulphates + alcohol
    )
s = summary(rlm)
s
rlm.d=data.frame(s$coefficients)
rlm.d$p.value =  2*pt(abs(rlm.d$t.value), s$df[2], lower.tail=FALSE) 
rlm.d
```
```{r echo=FALSE}
n=length(quality_score)
intercept = rep(1,n)
J=intercept%*%t(intercept) # get matrix of ones
I=diag(n)
X= cbind(intercept,type , volatile.acidity , citric.acid , residual.sugar , chlorides , free.sulfur.dioxide , total.sulfur.dioxide , sulphates , alcohol)
Y=quality_score.t
H= X%*%solve(t(X)%*%X)%*%t(X)
SSR=t(Y)%*%(H-J/n)%*%Y
SSTO=t(Y)%*%(I-J/n)%*%Y
SSE=SSTO-SSR
R_square=SSR/SSTO
R_adj=1-(SSE/(n-10))/(SSTO/(n-1))
c(R_square,R_adj)
```

# Diagnostic Plots of Reweight Regression Model
```{r echo=FALSE}
par(mfrow=c(2,2))
plot(rlm)
```

# Bootstrap
```{r warning=FALSE}
library(car)
set.seed(1000)
rlm.boot <- Boot(rlm, R=999)
summary(rlm.boot)
confint(rlm.boot,type = 'perc')
confint(best,level=0.95)
```

# Histgrams
```{r echo=FALSE}
#hist(rlm.boot,legend='separate')
```

# Result and Summary
